Cost- and Deadline-constrained Scheduling of Scientific Workflow Ensembles in IaaS Clouds

### 2012-01-20 paper/paper.tex
Changes to section 5.4

\begin{equation}
\label{eq:score}
s = \sum_{workflows\ completed}{2^{-d}}
\end{equation}

@@

We run the simulations for the constraint values ranging between the
estimated limits as indicated in Section~\ref{sec:scenarios}, resulting in
approximately 50-150 (budget, deadline) points for each application. The runs were
repeated for Pareto-distributed and constant workflow sizes. From this result
space we selected the most interesting plots which give the good overview  of
the observed algorithm behavior.

@@

Fig.~\ref{fig:number-complete-pareto} shows the number of the workflows
completed for two budget values per each application ensemble. The budget values
selected correspond to two opposite cases: {\em small budget} (top) and {\em
large budget} (bottom). The first one allows completing only a small number of
workflows from the ensemble, while the latter one suffices to finish almost all
the workflows. It can be seen that the static planning algorithm (SPSS) performs
better for tighter constraints: when the budget is low to admit only a small
number of workflows, the static procedure is able to more efficiently pack these
workflow tasks onto the resources. This can be clearly seen for the LIGO,
Epigenomics and SIPHT ensembles for small budgets and also for the same
ensembles with large budgets but short deadlines. The exception are Montage and
CyberShake ensembles, which consist of many fine-grained tasks that can be
executed in parallel. For these cases, as well for the other ensembles with
large budgets, the simple dynamic scheduling techniques perform equally well or
better. The reason for that behavior is that the large number of tasks
guarantees that resources are almost never idle, while the provisioner algorithm
makes sure that the VM instances are most efficiently utilized before shutdown.
We can also observe that the workflow-aware dynamic algorithm (WA-DPDS) performs
better than the workflow-unaware version. This means that for the tested
workflows the simple admission procedure based on estimation of workflow cost
does not degrade the solution, but rather in many cases it allows rejecting
larger workflows that would lead to budget overrun. Thus it can save the space
for smaller workflows that can complete.

@@

Fig.~\ref{fig:total-time} presents the same results using the total computing
time metric $T_C$. These plots confirm our previous observations about the
algorithms performance, but they also allow to observe a more general property.
For a given budget, when the deadline is tightened the number of work done
steeply decreases. Assigning a shorter deadline means that more VMs need to be
provisioned to complete the work, but this leads to higher parallelism and thus
to the lower parallel efficiency. E.g. running 2 VMs for 10 hours will be more
efficient than running 10 VMs for 2 hours. Therefore, if deadline permits and
costs are important, parallelism should be avoided when possible.

These observations are further confirmed when examining the effective resource
cost, as shown in Fig.~\ref{fig:cost}. It can be seen that lower resource
utilization and lower parallel efficiency for short deadlines lead to
substantial increase of actual costs of computing. This increase reaches
$\sim$20\%, which means that if we need to run e.g. a SIPHT workflow ensemble
with a short deadline, $\sim$\$200 out of \$1000 is spent on idle resources.

Another general conclusion can be drawn from Fig.~\ref{fig:cost}: despite some
random fluctuations resulting mostly from the finite result space, all the
curves for all the applications and budgets have a similar shape. This shape
represents a trade-off curve between two conflicting objectives: time and cost.
We can see that the selection of deadline and cost can be formulated as
multi-criteria optimization problem and the curves presented in
Fig.~\ref{fig:cost} are approximations of the Pareto front (or Pareto set) of
solutions to this problem. In our case the solutions were achieved by maximizing
the number of work done, which directly results in minimization of effective cost parameter for a given
deadline. Therefore, the obtained results can be used to assist the decision in
resource allocation planning, when both cost and time criteria are given not as
constraints but as objectives.

@@

In order to compare efficiency of algorithms, we counted the number of times
each algorithm achieved the best results for all tested applications, deadlines
and budgets. If two or more algorithms achieve the same highest results, all are
counted as highest. Table~\ref{tab:num-dags-pareto} shows these summary values
for Pareto-distributed workflows. These results depend on the metric: number of
workflows completed does not count the priorities or sizes of workflows, which
means that it happens more often that all the algorithms achieve the same (best)
result. However, it does not mean that all algorithms complete the same
workflows: that is why values in the tables (a), (b) and (c) are different.
Exponential score and total amount of computing time complete give the same
rankings, since in our test ensembles larger workflows were given higher
priorities. It can be observed that based on the sampled problem space the
workflow-aware algorithms achieve highest score in $\sim$2--3 times more cases
than the simple DPDS which does not take into account workflow structure or
size. In average, SPSS performs better in $\sim$20--50\% more times than
WA-DPDS. We have to note, that the sampled parameter space included the Montage
and CyberShake workflows consisting of many small tasks, for which cases SPSS
does not perform equally well.

>>>>

\begin{equation}
\label{eq:score}
s = \sum_{workflows\ completed}{2^{-p}}
\end{equation}

@@

We ran simulations for constraint values ranging between the estimated 
limits indicated in Section~\ref{sec:scenarios}, resulting in approximately 
50-150 (budget, deadline) data points for each application. The runs were
repeated for Pareto-distributed and constant workflow ensembles. From this result
space we selected the most interesting plots that characterize the observed 
behavior of the algorithms best.

@@

Fig.~\ref{fig:number-complete-pareto} shows the number of the workflows
completed for two budget values for each application ensemble. The budget values
selected correspond to two opposite cases: {\em small budget} (top) and {\em
large budget} (bottom). The small budgets permit the algorithms to complete 
only a small number of workflows from the ensemble, while the larger budgets 
enables the algorithms to finish almost all the workflows. In three of five 
cases the SPSS algorithm performs better for tighter constraints: when the 
budget is low, or the deadline is short, the static procedure is able to more 
efficiently pack workflow tasks onto limited resources. This is the case for 
the LIGO, Epigenomics and SIPHT ensembles for small budgets, and for large
budgets with short deadlines. The exceptions are the Montage and CyberShake 
ensembles,  which consist of many fine-grained tasks. In these applications, 
as well for the other ensembles with large budgets and deadlines, the simple 
dynamic scheduling techniques perform equally well or better. The reason for 
this behavior is that the large number of tasks guarantees that resources 
are almost never idle regardless of how the tasks are scheduled. We also 
observe that WA-DPDS performs better than the workflow-unaware version, DPDS. 
This means that for the tested workflows the simple admission procedure based 
on estimation of workflow cost does not degrade the solution, but rather in 
many cases it enables larger workflows that would lead to budget overrun to
be rejected. Thus it can save the space for smaller workflows to complete.

@@

Fig.~\ref{fig:total-time} presents the same results using the total computing
time metric, $T_C$. These plots confirm our previous observations about the
algorithms' performance, but they also allow us observe a more general property:
for a given budget, when the deadline is tightened the amount of work done
steeply decreases. Assigning a shorter deadline means that more VMs need to be
provisioned to complete the work, but this leads to higher parallelism and thus
to lower parallel efficiency. For example, running 2 VMs for 10 hours will tend
to be more efficient than running 10 VMs for 2 hours. Similarly, the only way 
to reduce cost when adding VMs is to achieve a super-linear speedup, which 
is unlikely in practice. Therefore, if the deadline permits, and cost is 
important, parallelism should be avoided when possible.

These observations are further confirmed when examining the effective computing
cost, as shown in Fig.~\ref{fig:cost}. It can be seen that lower resource
utilization and lower parallel efficiency for short deadlines lead to
substantial increases in the effective computing cost. This increase can be 
significant. For example, a SIPHT ensemble with short deadlines and a budget
of \$1,000 has an effective computing cost $\sim$8--20\% higher than the 
normal cost, depending on the algorithm used. That means that out of the 
\$1,000 budget, \$80--\$200 is spent on idle resources.

Another general conclusion can be drawn from Fig.~\ref{fig:cost}: despite some
random fluctuations resulting mostly from the finite result space, all the
curves for all the applications and budgets have a similar shape. This shape
represents a trade-off curve between two conflicting objectives: time and cost.
We observe that the selection of deadline and cost can be formulated as a
multi-criteria optimization problem, and that the curves presented in 
Fig.~\ref{fig:cost} are approximations of the Pareto front (or Pareto set) of
solutions to this problem. In our case the solutions were achieved by maximizing
the amount of work done, which directly results in minimization of effective 
computing cost metric for a given deadline. Therefore, the obtained results 
can be used to aid in resource allocation planning, when both cost and time 
criteria are given not as constraints but as objectives.

@@

In order to compare the efficiency of our algorithms, we counted the number of times
each algorithm achieved the best result for all simulated applications, deadlines
and budgets. Table~\ref{tab:num-dags-pareto} shows these summary values for 
Pareto-distributed workflows. Note that all algorithms that achieve the best 
result are credited, so the sum for a given application can be greater than the
total number of scenarios. These results depend on the metric: number of
workflows completed does not count the priorities or sizes of workflows, which
means that it happens more often that all the algorithms achieve the same best
result. However, it does not mean that all algorithms complete the same
workflows: that is why the values in (a), (b) and (c) are different. The 
exponential score and total amount of computing time completed give the same
rankings, since in our test ensembles larger workflows were given higher
priorities. It can be observed based on the sampled problem space that the
workflow-aware algorithms achieve highest score in $\sim$2--3 times more cases
than the simple DPDS algorithm which does not take into account workflow 
structure or size. SPSS performs the best in $\sim$20\% more cases than WA-DPDS,
including the Montage and CyberShake ensembles, for which SPSS does not 
perform as well as DPDS and WA-DPDS.
